%  $Id: caps_nap_menu.tex,v 1.8 2006/09/28 14:01:36 dav480 Exp $ 
    % Caps/Nap Menu

\section{Caps/Nap Menu}
    \label{caps-nap-menu}

\subsection{Introduction}
\label{caps-nap-menu-Introduction}

The distributed 
  \texttt{wish} startup file `\texttt{wishrc.tcl}'
    (which is copied to file `\texttt{.wishrc}' on unix systems)
    ends with the command
  `\texttt{caps\_nap\_menu}', which displays a GUI menu
  consisting of three buttons labelled 
  \emph{Browse}, 
  \emph{Command} and 
  \emph{Help}. This command `\texttt{caps\_nap\_menu}' is defined by the file
    `\texttt{caps\_nap\_menu.tcl}'. Each of these three buttons is
  used to display a sub-menu detailed in the following sections.

\subsection{Browse}
    \label{caps-nap-menu-Browse}

The 
  \emph{Browse} menu provides access to browsers for Tcl variables
  and a variety of files. These file browsers typically provide
  facilities to:
\begin{bullets}
    \item Select a file using the \texttt{choose\_file} GUI
	(see section \ref{choose-file}).
    \item Select data from this file.
    \item Display this data as text.
    \item Plot this data (as an XY graph or a 2D image) using 
    \texttt{plot\_nao} (see section \ref{plot-nao}).
    \item Print this data.
    \item Convert this data to a NAO.
\end{bullets}
Detailed help is available within each browser.

\subsubsection{Tcl Variables} 
\label{caps-nap-menu-Variables}

    This facility for browsing 
    \emph{Tcl variables} is defined by the file 
    \texttt{browse\_var.tcl}. It can be used to display the names
    and values of all tcl variables (including arrays), but has
    particular facilities for those referencing NAOs. The menu buttons
    have the following functions:
    \begin{description}
      \item[\emph{namespace}]
      Use tree widget to set the namespace.
      \item[\emph{list}]
      List tcl variables in the namespace matching the glob
      pattern. Display a line for each matching tcl variable. This line
      contains the variable's name and value provided there is
      room. Otherwise the line is truncated. You can click on a line to
      display:
    \begin{bullets}
        \item full value of an ordinary tcl variable or array whose line is truncated as above
        \item a NAO as either text or graph/image as specified by the radio-button
    \end{bullets}
      \item[\emph{help}] Display \emph{Help on Tcl Variable Browser}.
      \item[\emph{cancel}] Remove \emph{Tcl Variable Browser} widget.
    \end{description}

\subsubsection{\emph{AVHRR} and \emph{ATSR} Satellite Files}
\label{caps-nap-menu-AVHRR-ATSR}
	This item is only available if the Caps package is loaded.

\subsubsection{\emph{CIF} Files}
\label{caps-nap-menu-CIF}

    CIF files originated at Melbourne University. Their use has
    been largely replaced by netCDF but some data still exists in this
    format. This facility for browsing 
    \emph{CIF} files is defined by the file 
    \texttt{browse\_cif.tcl}.

\subsubsection{\emph{HDF} and \emph{netCDF} Files}
\label{caps-nap-menu-HDF-netCDF}

\paragraph{Introduction \\}
\label{caps-nap-menu-HDF-netCDF-intro}

HDF and netCDF are two similar common file formats used for
data in the form of arrays and grids, especially in the earth
sciences such as Meteorology and Oceanography.
See section \ref{hdf-netcdf} for further information about HDF 
and netCDF.

The term \emph{variable/SDS} is used below.
This means \emph{variable} for netCDF and \emph{SDS} for HDF.

The HDF/netCDF browser is a powerful tool which allows one to
select a variable/SDS or attribute from a displayed tree, then (for
variable/SDS) select a region within it.

The HDF/netCDF browser is defined by the file \texttt{hdf.tcl}.

\paragraph{Instructions}

\input{hdf.tex}

\subsubsection{\emph{Image} Files}
\label{caps-nap-menu-Image}

    This facility for 
    \emph{viewing image files} is defined by the file 
    \texttt{vif.tcl}. Standard Tk supports the PPM, PGM and GIF
    formats. If the package 
    \texttt{Img} is installed then it will be used, giving support
    for BMP, XBM, XPM, PNG, JPEG and TIFF as well. The ActiveTcl
    distribution of Tcl/Tk includes 
    \texttt{Img}.

\subsection{Command}
    \label{caps-nap-menu-Command}

The 
  \emph{Command} menu allows the following common Tcl commands to be
  executed with a mouse click:
  \begin{description}
    \item[\label{caps-nap-menu-History} \texttt{history}]
    Display command history. This differs from the command-line 
    \texttt{history} command in that it uses a scrolled window.
    Note that 
    \texttt{tkcon} also provides:
    \begin{bullets}
      \item \emph{History} entry in main menu
      \item \emph{History} entry within \emph{File}/\emph{Save}
    \end{bullets}
    \item[\label{caps-nap-menu-Exit} \texttt{exit}] Quit.
  \end{description}

  \subsection{Help}
    \label{caps-nap-menu-Help}

The 
  \emph{Help} menu provides access to local and Web documentation on:
\begin{bullets}
    \item Caps/Nap Menu
    \item Tcl/Tk
    \item Nap
    \item Caps
\end{bullets}
Note that the Web versions may be more recent than the installed
  local documentation.
