%  $Id: model.tex,v 1.7 2006/06/07 04:58:19 dav480 Exp $ 
    % NAP: Data Models

\section{Data Models}

A 
  \textit{data model} is a mental model of the nature of some data. It
  answers such questions as the following:

\subsection{What values can the data take?}

Are they all numeric? Are they
  all integers? Is the set of possible values finite? What are the
  minimum and maximum possible values? Are $\infty$ and 
  \textit{NaN} possible values? Do some values have special meanings,
  such as indicating undefined or missing data?

\subsection{What is the measurement level?}

Data is often classified as follows according to \emph{measurement level}:
\\
    \begin{tabular}{|l||p{40 mm}|p{20 mm}|p{20 mm}|p{30 mm}|}
    \hline 
      \textbf{Level} & 
      \textbf{Description} & 
      \textbf{Valid \mbox{Operations}} & 
      \textbf{Measure of Central\ Tendency} & 
      \textbf{Examples}
    \\
    \hline 
    \hline 
      nominal & 
      Values denote categories which have no order & 
      = $\neq$ & 
      mode & 
      \mbox{zip post-code} \mbox{chemical e.g. CO$_{2}$}
    \\
    \hline 
      ordinal & 
      \mbox{Values ordered} \mbox{Differences meaningless} & 
      = $\neq$ $<$ $\le$ $>$ $\ge$ & 
      median & 
      Richter earthquake scale
    \\
    \hline 
      interval & 
      \mbox{Differences valid} \mbox{Quotients meaningless} & 
      $= \neq < \le > \ge + -$ & 
      arithmetic mean & 
      temperature in $^{\circ}$C
    \\
    \hline 
      ratio & 
      Quotients valid & 
      $= \neq < \le > \ge + - \times \div $ & 
      geometric mean & 
      temperature in $^{\circ}$K 
    \\
  \hline
\end{tabular}

\subsection{How accurate are the values?}

A \textit{measurement error} is the difference between the 
  \textit{true value} and the 
  \textit{measured value}. Measured values can differ from true values
  due to:
\begin{bullets}
    \item finite precision of instrument
    \item systematic errors (e.g. inadequately calibrated
    instrument)
    \item random errors (due to finite size of sample)
    \item sampling errors (due to non-randomness of sample)
    \item blunders (e.g. a human misreading an instrument)
\end{bullets}
It is desirable to include error estimates with data.

\subsection{Are the data located in some space?}

A \textit{time series} consists of values located along the time
  dimension. 
  \textit{Geographic} data is located along spatial dimensions such as
  latitude, longitude and altitude and may also have a time dimension.
  Note that longitude is 
  \textit{cyclic}.

The dimensions of the space can have a measurement level of \textit{nominal}.
For example, an accounting spreadsheet might have columns corresponding to 
\textit{charge codes} and rows corresponding to \textit{company divisions}.

Some variables have a vector value at each point in space and time.
For example, wind data is typically stored as two (east and north) components. 
These components are often treated as separate variables.
An alternative approach is to have a single array whose dimensions include a
\textit{nominal} dimension corresponding to the components.
Thus a wind array could include a dimension of size 2 corresponding to east and north.

Data located in a continuous space can be either 
  \textit{gridded} or 
  \textit{scattered}. Both types are discussed in section \ref{grid} (Nap Grids).
