%  $Id: nap_get.tex,v 1.7 2006/09/28 05:12:30 dav480 Exp $ 
    % nap\_get

\label{nap-get}

\section{Introduction}
    \label{nap-get-Introduction}

The 
  \texttt{nap\_get} command creates a NAO containing data read from
  a file. The first argument specifies the type of file, which can be 
  \texttt{binary}, 
  \texttt{hdf}, 
  \texttt{netcdf} or 
  \texttt{swap}.

The HDF and netCDF file formats are discussed in section \ref{hdf-netcdf}.
The \texttt{netcdf} option supports (but only on the Linux Intel
  386 platform at the time of writing) access to remote data provided
  by an 
  \href{http://www.opendap.org/}{OPeNDAP (a.k.a. DODS)} server.
  In this case the 
  $\mathit{filename}$ is a URL rather than the pathname of a local
  file. There is an OPeNDAP example at the end of 
  section \ref{nap-get-Reading-netCDF-Data} (\emph{Reading netCDF Data}).

\section{Reading Binary Data}
    \label{nap-get-Reading-Binary-Data}

Binary data is read using the command
\\
\texttt{nap\_get binary} $\mathit{channel}$ [$\mathit{datatype}$ [$\mathit{shape}$]]
\\
where 
  $\mathit{datatype}$ defaults to 
  \texttt{u8} and 
  $\mathit{shape}$ defaults to a vector containing a single element, the file size.

The command
\\
\texttt{nap\_get swap} $\mathit{channel}$ [$\mathit{datatype}$ [$\mathit{shape}$]]
\\
is similar, except that bytes are swapped.
This enables reading of data written on a machine with opposite byte-order within words.

The following example first writes six 32-bit floating-point
  values to a file using the OOC 
  \texttt{binary} method. It then reads them back into a NAO named
  `\texttt{in}' using `\texttt{nap\_get binary}':
  \begin{verbatim}
% set file [open float.dat w]
filee15170
% [nap "f32{1.5 -3 0 2 4 5}"] binary $file
% close $file
% set file [open float.dat]
filee15170
% nap "in = [nap_get binary $file f32]"
::NAP::22-22
% close $file
% $in all
::NAP::22-22  f32  MissingValue: NaN  References: 1
Dimension 0   Size: 6      Name: (NULL)    Coordinate-variable: (NULL)
Value:
1.5 -3 0 2 4 5
\end{verbatim}

Note that no shape was specified, giving a 6-element vector. The
following example reads the file again, this time specifying a shape of
  \texttt{\{2 3\}}. The NAO is displayed but not saved.
  \begin{verbatim}
% set file [open float.dat]
file6
% [nap_get binary $file f32 "{2 3}"] all
::NAP::32-32  f32  MissingValue: NaN  References: 0  Unit: (NULL)
Dimension 0   Size: 2      Name: (NULL)    Coordinate-variable: (NULL)
Dimension 1   Size: 3      Name: (NULL)    Coordinate-variable: (NULL)
Value:
 1.5 -3.0  0.0
 2.0  4.0  5.0
% close $file
\end{verbatim}

\section{Reading netCDF Data}
    \label{nap-get-Reading-netCDF-Data}

NetCDF data is read using the command
  \\
  \texttt{nap\_get netcdf} $\mathit{filename}$ $\mathit{name}$ [$\mathit{index}$ [$\mathit{raw}$]]

\noindent
  $\mathit{filename}$ is either:
\begin{bullets}
    \item pathname of a local file
    \item URL of a remote 
    \href{http://www.opendap.org/}{OPeNDAP}
    server
\end{bullets}

\noindent
$\mathit{name}$ is the name of a variable or attribute and has one of
the following forms:
\begin{bullets}
    \item $\mathit{varname}$ for a variable
    \item $\mathit{varname}\texttt{:}\mathit{attribute}$ for an attribute of a variable
    \item $\texttt{:}\mathit{attribute}$ for a global attribute
\end{bullets}

A single-element attribute gives a scalar. Other attributes give
  vectors. Neither 
  $\mathit{index}$ nor 
  $\mathit{raw}$ can be specified for attributes.

A variable with no 
  $\mathit{index}$ gives the entire variable.
If $\mathit{index}$ is specified then it selects using 
{\em cross-product} indexing, as described in section
  \ref{indexing-Cross-product-index}.

If 
  $\mathit{raw}$ is 1 then the result contains raw data read from the
  file. If 
  $\mathit{raw}$ is 0 (default) then this data is transformed using the
  attributes 
  \texttt{scale\_factor}, 
  \texttt{add\_offset}, 
  \texttt{\_FillValue}, 
  \texttt{valid\_min}, 
  \texttt{valid\_max} and 
  \texttt{valid\_range} if any of these are present.
  

The following example first creates a netCDF file using the netCDF
  utility 
  \texttt{ncgen}. There is one variable called 
  \texttt{vec}. It is a 3-element 32-bit integer vector with
  elements 6, -9 and 4. The data is read into a NAO called 
  \texttt{v} using 
  \texttt{nap\_get netcdf}.
  \begin{verbatim}
% exec ncgen -b << {
    netcdf int {
        dimensions:
            n = 3 ;
        variables:
            int vec(n) ;
        data:
            vec = 6, -9, 4 ;
    }
}
% nap "v = [nap_get netcdf int.nc vec]"
::NAP::52-52
% $v all
::NAP::52-52  i32  MissingValue: -2147483648  References: 1  Unit: (NULL)
Dimension 0   Size: 3      Name: n         Coordinate-variable: (NULL)
Value:
6 -9 4
\end{verbatim}

  

Now let's specify the index 
  \texttt{\{0 2\}} to select the first and third elements:
  \begin{verbatim}
% [nap_get netcdf int.nc vec "{0 2}"] all
::NAP::58-58  i32  MissingValue: -2147483648  References: 0  Unit: (NULL)
Dimension 0   Size: 2      Name: (NULL)    Coordinate-variable: (NULL)
Value:
6 4
\end{verbatim}

  

The following shows the different effects of a single-element
  vector index and a scalar index with the same value:
  \begin{verbatim}
% [nap_get netcdf int.nc vec "{1}"] all
::NAP::65-65  i32  MissingValue: -2147483648  References: 0
Dimension 0   Size: 1      Name: n         Coordinate-variable: ::NAP::69-69
Value:
-9
% [nap_get netcdf int.nc vec "1"] all
::NAP::83-83  i32  MissingValue: -2147483648  References: 0
Value:
-9
\end{verbatim}

The following is an 
  \href{http://www.opendap.org/}{OPeNDAP (a.k.a. DODS)} example.
  It displays the altitude of the south pole.
  \begin{verbatim}
% set url \
http://www.marine.csiro.au/dods/nph-dods/dods-data/climatology-netcdf/etopo5.nc
% [nap_get netcdf $url height "@@-90, 0"]
2810
\end{verbatim}

\section{Reading HDF Data}
    \label{nap-get-Reading-HDF-Data}

HDF data is read using the command
\\
\texttt{nap\_get hdf} $\mathit{filename}$ $\mathit{name}$ [$\mathit{index}$ [$\mathit{raw}$]]

\noindent
  $\mathit{filename}$ is the pathname of a local file.

\noindent
  $\mathit{name}$ is the name of an SDS or attribute and has 
one of the following forms:
\begin{bullets}
    \item $\mathit{sdsname}$ for a SDS
    \item $\mathit{sdsname}$\texttt{:}$\mathit{attribute}$ for an attribute of a SDS
    \item \texttt{:}$\mathit{attribute}$ for a global attribute
\end{bullets}
  
A single-element attribute gives a scalar. Other attributes give
  vectors. Neither 
  $\mathit{index}$ nor 
  $\mathit{raw}$ can be specified for attributes.

An SDS with no $\mathit{index}$ gives the entire SDS.
If $\mathit{index}$ is specified then it selects using 
{\em cross-product} indexing, as described in section
  \ref{indexing-Cross-product-index}.

If 
  $\mathit{raw}$ is 1 then the result contains raw data read from the
  file. If 
  $\mathit{raw}$ is 0 (default) then this data is transformed using the
  attributes 
  \texttt{scale\_factor}, 
  \texttt{add\_offset}, 
  \texttt{\_FillValue}, 
  \texttt{valid\_min}, 
  \texttt{valid\_max} and 
  \texttt{valid\_range} if any of these are present.
  
The following example writes data to an HDF file using the OOC 
  \texttt{hdf} method. Then 
  \texttt{nap\_get hdf} is used with various index values (including
  default) to read the data back into temporary NAOs (which are deleted
  after being displayed):
  \begin{verbatim}
% [nap "f64{{1 0 9}{3 2 -1}}"] hdf mat.hdf mat64
% [nap_get hdf mat.hdf mat64] all; # default index giving whole SDS
::NAP::47-47  f64  MissingValue: NaN  References: 0
Dimension 0   Size: 2      Name: fakeDim0  Coordinate-variable: (NULL)
Dimension 1   Size: 3      Name: fakeDim1  Coordinate-variable: (NULL)
Value:
 1  0  9
 3  2 -1
% [nap_get hdf mat.hdf mat64 "1,0"] all; # select element
giving scalar
::NAP::78-78  f64  MissingValue: NaN  References: 0
Value:
3
% [nap_get hdf mat.hdf mat64 "{1},0"] all; # select element
giving vector
::NAP::102-102  f64  MissingValue: NaN  References: 0
Dimension 0   Size: 1      Name: fakeDim0  Coordinate-variable: ::NAP::92-92
Value:
3
% [nap_get hdf mat.hdf mat64 "0,"] all; # select row
::NAP::123-123  f64  MissingValue: NaN  References: 0
Dimension 0   Size: 3      Name: fakeDim1  Coordinate-variable: (NULL)
Value:
1 0 9
% [nap_get hdf mat.hdf mat64 ",2"] all; # select column
::NAP::147-147  f64  MissingValue: NaN  References: 0
Dimension 0   Size: 2      Name: fakeDim0  Coordinate-variable: (NULL)
Value:
9 -1
% [nap_get hdf mat.hdf mat64 ",{0 2}"] all; # select
sub-matrix
::NAP::154-154  f64  MissingValue: NaN  References: 0
Dimension 0   Size: 2      Name: fakeDim0  Coordinate-variable: (NULL)
Dimension 1   Size: 2      Name: fakeDim1  Coordinate-variable: ::NAP::161-161
Value:
 1  9
 3 -1
\end{verbatim}

\section{Listing Names of Variables/SDSs and Attributes in HDF and netCDF Files}
    \label{nap-get-Listing-Names}

One can list the names of variables/SDSs and attributes matching
  a regular expression 
  $\mathit{RE}$ using the command
\\
\texttt{nap\_get hdf -list} $\mathit{filename}$ [$\mathit{RE}$]
\\or
\\
\texttt{nap\_get netcdf -list} $\mathit{filename}$ [$\mathit{RE}$]
\\All variables/SDSs and attributes are listed if 
  $\mathit{RE}$ is omitted. For example, using the HDF file created
  above:
  \begin{verbatim}
% nap_get hdf -list mat.hdf
mat64
mat64:_FillValue
\end{verbatim}

Some useful regular expressions are

\begin{tabular}{|l|l|}
    \hline 
      \textbf{Regular Expression} & \textbf{Select all:}
    \\
    \hline 
    \hline 
        \verb/^[^:]*$/ & variables/SDSs
    \\
    \hline 
        \verb/:/ & attributes
    \\
    \hline 
        \verb/^:/ & global attributes
    \\
    \hline 
        \verb/.:/ & non-global attributes
    \\
  \hline
\end{tabular}

\vspace{1 mm}

Thus we can restrict the above list to SDSs only using:
  \begin{verbatim}
% nap_get hdf -list mat.hdf {^[^:]*$}
mat64
\end{verbatim}

\section{Reading Metadata from HDF and netCDF Files}
    \label{nap-get-Metadata}

The command
  \\
  \texttt{nap\_get hdf -datatype} 
  $\mathit{filename}$ $\mathit{sdsname}$
  \\or
  \\
  \texttt{nap\_get netcdf -datatype} 
  $\mathit{filename}$ $\mathit{varname}$
  \\returns the data-type of a specified SDS/variable in the
  specified file.
  

The command
  \\
  \texttt{nap\_get hdf -rank} 
  $\mathit{filename}$ $\mathit{sdsname}$
  \\or
  \\
  \texttt{nap\_get netcdf -rank} 
  $\mathit{filename}$ $\mathit{varname}$
  \\returns the rank (number of dimensions) of a specified
  SDS/variable in the specified file.
  

The command
  \\
  \texttt{nap\_get hdf -shape} 
  $\mathit{filename}$ $\mathit{sdsname}$
  \\or
  \\
  \texttt{nap\_get netcdf -shape} 
  $\mathit{filename}$ $\mathit{varname}$
  \\returns the shape (dimension sizes) of a specified SDS/variable
  in the specified file.
  

The command
  \\
  \texttt{nap\_get hdf -dimension} 
  $\mathit{filename}$ $\mathit{sdsname}$
  \\or
  \\
  \texttt{nap\_get netcdf -dimension} 
  $\mathit{filename}$ $\mathit{varname}$
  \\returns the dimension names of a specified SDS/variable in the
  specified file.
  

The command
  \\
  \texttt{nap\_get hdf -coordinate} $\mathit{filename}$ $\mathit{sdsname}$ $\mathit{dim\_name}|\mathit{dim\_number}$
  \\or
  \\
  \texttt{nap\_get netcdf -coordinate} $\mathit{filename}$ $\mathit{\mathit{varname}}$ $\mathit{\mathit{dim\_name}}|\mathit{\mathit{dim\_number}}$
  \\returns the coordinate variable NAO corresponding to the
  specified dimension of the specified SDS/variable in the specified
  file.
