%  $Id: refs.tex,v 1.5 2007/01/01 23:31:26 dav480 Exp $ 
    % References

\section{References}

\subsection{Introduction}
    \label{refs-Introduction}

The following provides a guide to books, papers and web sites
  relevant to Nap. The opinions are those of the author (Harvey
  Davies).

\subsection{Tcl/Tk}
    \label{refs-tcl}

The following are my three favourite Tcl/Tk web sites:
  \begin{description}
    \item[
      \href{http://wiki.tcl.tk}{wiki.tcl.tk}
    ]
    Wiki-based open community site. Excellent starting point.
    \item[
      \href{http://www.tcl.tk}{www.tcl.tk}
    ]
    Main Tcl Developer Xchange site.
    \item[
      \href{http://www.activestate.com/Products/ActiveTcl/}{www.activestate.com/Products/ActiveTcl/}
    ]
    ActiveState provide ActiveTcl here for downloading.
  \end{description}
  \par Despite its age, I consider John Ousterhout's 1994 
  \emph{Tcl and the Tk Toolkit} to still be the best introduction
  to Tcl/Tk. In fact it really is a classic example of clear readable
  technical writing. Ousterhout was the original developer of Tcl. Tcl
  has developed a lot since 1994 and the new features missing from this
  book are discussed in the 
  \href{http://wiki.tcl.tk/103}{Wiki}.
  \par The only detailed up-to-date book is the much less readable 4th
  edition of 
  \emph{Practical Programming in Tcl and Tk} by Brent Welch and
  Ken Jones with Jeffrey Hobbs. This book has the web site 
  \href{http://www.beedub.com/book}{www.beedub.com/book}.

\subsection{Nap}
    \label{refs-Nap}

The home page for 
  \emph{nap} (a.k.a.
  \emph{tcl-nap}) is at 
  \href{http://tcl-nap.sourceforge.net/index.php}{http://tcl-nap.sourceforge.net/index.php}.
The main documentation of Nap is provided by the
  \href{http://tcl-nap.sourceforge.net/nap_users_guide.pdf}{Nap User's Guide},
which is what you are reading.
  \par The only published paper on Nap is Harvey Davies's talk 
  \emph{The NAP (N-Dimensional Array Processor) Extension to Tcl}
  , which was given at the 9th (2002) Annual Tcl/Tk Conference.
Both the
\href{http://aspn.activestate.com/ASPN/Tcl/TclConferencePapers2002/Tcl2002papers/davies-nap/nap.pdf}{original}
and a
  \href{http://tcl-nap.sourceforge.net/nap_paper2002.pdf}{revised} version are available.
  \par Nap was originally developed as part of the CSIRO Caps project,
  which has the web page 
  \href{http://www.eoc.csiro.au/cats/caps/}{http://www.eoc.csiro.au/cats/caps/}.

\subsection{Array Processing}
    \label{refs-array}

There are two main array processing professional associations in
  the world. The USA has the ACM 
  \emph{Special Interest Group on the APL and J languages}, which has
  the web page 
  \href{http://www.acm.org/sigapl}{www.acm.org/sigapl}. The
  British Computer Society has a specialist group the 
  \emph{British APL Association}, which publishes the journal 
  \emph{Vector}, which has the web page 
  \href{http://www.vector.org.uk}{www.vector.org.uk}. Both
  associations were originally concerned with only APL but now include
  other array processing languages.
  \par The primary site for the J language is 
  \href{http://www.jsoftware.com}{www.jsoftware.com}.

\subsection{Array-oriented File Formats}
    \label{refs-file}

The primary site for netCDF is 
  \href{http://www.unidata.ucar.edu/packages/netcdf/index.html}{www.unidata.ucar.edu/packages/netcdf/index.html}.
  \\The primary site for HDF is 
  \href{http://www.hdfgroup.org/}{www.hdfgroup.org/}.

\subsection{Arithmetic}
    \label{refs-Arithmetic}

The classic article on floating-point arithmetic (and the IEEE
  Standard 754 for it) is David Goldberg's 
\href{http://portal.acm.org/citation.cfm?id=103163&coll=GUIDE&dl=ACM&CFID=36776917&CFTOKEN=647870&ret=1#Fulltext}
{\emph{What Every Computer Scientist Should Know About Floating-Point Arithmetic}},
ACM Computing Surveys, 1991.
A more recent paper is Kahan's
  \href{http://www.cs.berkeley.edu/~wkahan/ieee754status/ieee754.ps} {\emph{Lecture 
    Notes on the Status of IEEE 754} }.
