%  $Id: scripts.tex,v 1.6 2006/06/07 04:58:22 dav480 Exp $ 
    % Sample Startup Scripts

\section{Sample Startup Scripts}
    \label{scripts}
    \label{scripts-Sample-Startup-Scripts}

\subsection{Introduction}
    \label{scripts-Introduction}

These startup scripts tailor 
  \emph{Tcl} and 
  \emph{tkcon} to personal requirements. The scripts can be copied to
  your home directory and modified as desired.

\subsection{Script \texttt{my.tcl}}
    \label{scripts-my}

The script 
  \verb!~/my.tcl! is called by the script 
  \verb!~/.tclshrc! or 
  \verb!~/tclshrc.tcl! that is created by the standard nap
  installation. This 
  \verb!~/my.tcl! script can be used to tailor Tcl in various
  ways such as
\begin{bullets}
    \item initialising packages
    \item initialising libraries
    \item defining procedures
    \item sourcing other scripts
\end{bullets}
  
 The following sample script initialises a personal library and
  defines some simple aliases:
  \begin{verbatim}
# my.tcl --
#
# Initialise my library of Tcl procedures (including Nap functions) by
sourcing its tclIndex.
# The library consists of files in directory ~/my_tcl_lib
# The file tclIndex is created by the following command:
#     auto_mkindex ~/my_tcl_lib *.tcl
foreach dir {~/my_tcl_lib} {source $dir/tclIndex}
#
proc define_alias {alias args} {proc $alias args "eval $args \$args"}
#
define_alias e expr
define_alias n nap
define_alias p pwd
define_alias cp file copy
define_alias md file mkdir
define_alias rm file delete
\end{verbatim}

\subsection{Script \texttt{my\_tkcon.tcl}}
    \label{scripts-my-tkcon}

The script 
  \verb!~/my_tkcon.tcl! is called by the script 
  \verb!~/.tkconrc! or 
  \verb!~/tkcon.cfg! that is created by the standard nap
  installation. This 
  \verb!~/my_tkcon.tcl! script can be used to tailor various
  aspects of tkcon.
  
 The following sample script defines
\begin{bullets}
    \item the font used by tkcon
    \item the number of rows and columns provided by tkcon
    \item the number of commands in history
    \item the prompt at start of each input line
\end{bullets}
\begin{verbatim}
# my_tkcon.cfg --
#
set ::tkcon::OPT(font) "Courier 10"
set ::tkcon::OPT(rows) 40
set ::tkcon::OPT(cols) 80
set ::tkcon::OPT(history) 200; # number of commands in history
set ::tkcon::OPT(prompt1) {% };# prompt at start of each input line
# set ::tkcon::OPT(prompt1) {[history nextid] % };# alternative
\end{verbatim}

